Complex Analysis of Askaryan Radiation: UHE-$\nu$ Identification and Reconstruction via the Hilbert Envelope of Observed Signals

### 2025-10-07 template.tex
minor changes ... time to print

To reduce uncertainties, the Hilbert envelope of observed data is used in cross-correlations instead of the original signals.  Thus, an analytic prediction for the Hilbert envelope of the observed data would an effective signal template.  An assumption must be made, however, for the RF channel response.  The RLC damped oscillator is a suitable circuit model for the RF dipole antennas used in RICE, RNO-G, and the proposed IceCube Gen2 \cite{10.1088/1748-0221/16/03/p03025,10.48550/arxiv.2008.04323}.  In fact, an RLC response was first used by RICE at the South Pole a decade ago \cite{10.1103/PhysRevD.85.062004}.

@@

The detector was \textit{a single string} of 8 RF dipoles, each with the same (x,y)-coordinates and regularly-spaced z-coordinates.  The ice had a depth and radius of 0.65 km and 0.85 km, respectively, and the dielectric properties of the South Pole.  Each RF channel had a filtered, amplified passband of [0.08-1] GHz, 256 samples, and a 1 GHz sampling rate.  The RF trigger responded when any 3 of the 8 voltage traces exceeded $\pm 3$ times the rms voltage ($v_{\rm rms}$) of the thermal noise (233K) within 256 ns.  The Hilbert envelope of the coherently summed waveform (CSW) was calculated from the traces and cross-correlated with Eqs. \ref{eq:final}-\ref{eq:final2}.  Each correlation was maximized by varying $\sigma_t$.  An optimization using noiseless data (3K) yielded 0.15 GHz and 0.025 GHz for $f_0$ and $\gamma$, respectively.  For each signal, a thermal noise trace satisfying the RF trigger was generated and cross-correlated against the best-fit analytic envelope.  The results are shown in Fig. \ref{fig:fig3}.

\begin{table}
\small
\centering
\begin{tabular}{| c | c | c |}
\hline
\textbf{Parameter} & \textbf{Value} & \textbf{Note} \\
Ice Model & South Pole & 2015 measurements \\
Signal Model & ARZ2020 & (see \cite{PhysRevD.101.083005}) \\
Trigger & 3 of 8 channels & high-low trigger, $3v_{\rm rms}$ \\
RF channels & 8 & RF bicone (in firn) \\
Channel filters & [80-1000] MHz & Passband \\
Noise Temperature & 233K & Sets $v_{\rm rms}$ \\
Sampling Rate & 1 GHz & $f_{\rm c} = 500$ MHz \\
Samples per channel & 256 &  \\
Channel depths & [-4,-6,-8,...-18] m & cable delays included \\
RF cable type & LMR-400 & $<0.0886$ dB m$^{-1}$ \\
\hline
\end{tabular}
\caption{\label{tab:1} Important NuRadioMC parameters used in this analysis.}
\end{table}

@@

In Secs. \ref{sec:int} and ref{sec:unit}, previous work in the utility of analytic Askaryan models was reviewed, and important parameters were defined.  In Sec. \ref{sec:onc}, a fully analytic model for $\mathcal{E}_{\rm s*r}(t)$ and $s(t) * r(t)$ was presented and checked for accuracy.  In Sec. \ref{sec:sim}, the analytic model was shown to reject all but 0.2 RF thermal triggers in a 5 year run at a 1 Hz thermal trigger rate using a correlation threshold of 0.4.  That same threshold preserves 99.99\% of UHE-$\nu$ signals at 100 PeV.  The correlation distribution for UHE-$\nu$ signals peaked at 0.94, and some values even exceed 0.94.  Finally, in Sec. \ref{sec:recon}, a first attempt at using the CSW fit with data \textit{a single-string detector} to reconstruct $\log_{10}E_{\rm C}$ was presented.  Though the error range for $\log_{10}E_{\rm C}$ is large, the average value is correct: $\log_{10}E_{\rm C} = 16.5 \pm 2.0$ for 100 PeV.

@@

We would like to thank Prof. Steve Barwick, and Prof. Mark Kozek for useful background discussions.  We are grateful to the authors of RB, Prof. John Ralston and Roman Buniy, for inspiring us in this line of research \cite{10.1103/physrevd.65.016003}.  We would like to thank Prof. Dave Besson for encouragement and wisdom while we continued to develop the model.  Most of all, we would like to thank Prof. Amy Connolly for inspiring us and encouraging us to master more advanced applications of complex analysis to mathematical physics.  It is our hope that the model will one day help to identify the first UHE-$\nu$ measured via the Askaryan effect.

>>>>

To reduce uncertainties, the Hilbert envelope of observed data is used in cross-correlations instead of the original signals.  Thus, an analytic prediction for the Hilbert envelope of the observed data would an effective signal template.  An assumption must be made, however, for the RF channel response.  The RLC damped oscillator is a suitable circuit model for the RF dipole antennas used in RICE, RNO-G, and the proposed IceCube Gen2 \cite{10.1088/1748-0221/16/03/p03025,10.48550/arxiv.2008.04323}.  In fact, an RLC response was first used by RICE at the South Pole \cite{10.1103/PhysRevD.85.062004}.

@@

The detector was \textit{a single string} of 8 RF dipoles, each with the same (x,y)-coordinates and regularly-spaced z-coordinates.  The ice had a depth and radius of 0.65 km and 0.85 km, respectively, and the dielectric properties of the South Pole.  Each RF channel had a filtered, amplified passband of [0.08-1] GHz, 256 samples, and a 1 GHz sampling rate.  The RF trigger responded when any 3 of the 8 voltage traces exceeded $+3$ times \textit{and} $-3$ times the rms voltage ($v_{\rm rms}$) of the thermal noise ($T=233$ K) within 256 ns (256 samples).  The Hilbert envelope of the coherently summed waveform (CSW) was calculated from the traces and cross-correlated with Eqs. \ref{eq:final}-\ref{eq:final2}.  Each correlation was maximized by varying $\sigma_t$.  An optimization using noiseless data (3K) yielded 0.15 GHz and 0.025 GHz for $f_0$ and $\gamma$, respectively.  For each signal, a thermal noise trace satisfying the RF trigger was generated and cross-correlated against the best-fit analytic envelope.  The results are shown in Fig. \ref{fig:fig3}.

\begin{table}
\small
\centering
\begin{tabular}{| c | c | c |}
\hline
\textbf{Parameter} & \textbf{Value} & \textbf{Note} \\
Ice Model & South Pole & 2015 measurements \\
Signal Model & ARZ2020 & (see \cite{PhysRevD.101.083005}) \\
Trigger & 3 of 8 channels & $\pm 3v_{\rm rms}$ \\
RF channels & 8 & RF bicone (in firn) \\
Channel filters & [80-1000] MHz & Passband \\
Noise Temperature & 233K & Sets $v_{\rm rms}$ \\
Sampling Rate & 1 GHz & $f_{\rm c} = 500$ MHz \\
Samples per channel & 256 & total time, $256$ ns \\
Channel depths & [-4,-6,-8,...-18] m & cable delays included \\
RF cable type & LMR-400 & $<0.0886$ dB m$^{-1}$ \\
\hline
\end{tabular}
\caption{\label{tab:1} Important NuRadioMC parameters used in this analysis.}
\end{table}

@@

In Secs. \ref{sec:int} and \ref{sec:unit}, previous work in the utility of analytic Askaryan models was reviewed, and important parameters were defined.  In Sec. \ref{sec:onc}, a fully analytic model for $\mathcal{E}_{\rm s*r}(t)$ and $s(t) * r(t)$ was presented and checked for accuracy.  In Sec. \ref{sec:sim}, the analytic model was shown to reject all but 0.2 RF thermal triggers in a 5 year run at a 1 Hz thermal trigger rate using a correlation threshold of $\rho\geq 0.4$.  The threshold $\rho\geq 0.4$ preserves 99.99\% of UHE-$\nu$ signals at 100 PeV.  The correlation distribution for UHE-$\nu$ signals peaked at 0.94, and some values even exceed 0.94.  Finally, in Sec. \ref{sec:recon}, a first attempt at using the CSW fit with data \textit{a single-string detector} to reconstruct $\log_{10}E_{\rm C}$ was presented.  Though the error range for $\log_{10}E_{\rm C}$ was large, the average value was correct: $\log_{10}E_{\rm C} = 16.5 \pm 2.0$ for 100 PeV.

@@

We would like to thank Prof. Steve Barwick, and Prof. Mark Kozek for useful background discussions.  We are grateful to the authors of RB, Prof. John Ralston and Roman Buniy, for inspiring us in this line of research \cite{10.1103/physrevd.65.016003}.  We would like to thank Prof. Dave Besson for encouragement and wisdom while we continued to develop the model.  Most of all, we would like to thank Prof. Amy Connolly for inspiring us and encouraging us to master more advanced applications of complex analysis to mathematical physics.  We hope the model will help identify the first UHE-$\nu$ via the Askaryan effect, especially in light of potential UHE-$\nu$ observations \cite{collaboration2025observation-22f}.
